---
title: "emprac-stats"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{emprac-stats}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This vignette illustrates the use of the dataset included in the package 'emprac'. I initially collected this dataset to analyze the average vocal pitch-class imprecision and inaccuracy in 10 global musical regions, so I will demonstrate how I statistically analyzed some of these data and plotted the results.

## Preliminaries
Load the packages used in this vignette:

```{r setup, results='hide', warning=FALSE, message=FALSE}
r = getOption("repos")
r["CRAN"] = "http://cran.us.r-project.org"
options(repos = r)
install.packages("devtools")
devtools::install_github("phille10/emprac")
library(emprac)
library(ggplot2)
library(gridExtra)
library(RColorBrewer)
library(tidyverse)
library(broom)
library(AICcmodavg)
library(pwr)
library(sandwich)
library(multcomp)
#library(leveneTest)
library(car)
library(moments)
options(contrast=c("contr.sum","contr.poly"))

```

## Loading the data
To load the data, use the function 'data()': 

```{r load-data}
data("pitchclass_range", package="emprac")
```

To inspect the dataframe, use the function 'summary()':

```{r summary-data}
summary(pitchclass_range)
```

The data frame consists of 7 variables. The variable definitions can be consulted in the help file:

```{r call-help}
?pitchclass_range
```

## Examining pitch-class imprecision across regions
The region labels  are currently just characters, and we'd like them to be categorical variables (in R, levels of a factor). Let's do that, and combine the region data with just the range data to create a cleaner dataframe for analysis:
```{r range-cleanup}
#Define region levels
regionlevels <- c('Africa','Australia','Central Asia','Circumpolar','East Asia','Europe','Insular Pacific','Middle East North Africa','North America','South America')
#Create abbreviated dataframe
regions <- pitchclass_range$region.category
ranges <- pitchclass_range$pitch.class.range
rangedata <- data.frame(regions, ranges)
#Make region a categorical variable
rangedata$regions <- factor(rangedata$regions, levels=c('Africa','Australia','Central Asia','Circumpolar','East Asia','Europe','Insular Pacific','Middle East North Africa','North America','South America'))
#get rid of computational outliers
rangedata <- subset(rangedata, ranges < 400)
#Re-examine the data
summary(rangedata)
```

Now, let's run a simple linear regression on the data:
```{r lm.01}
lm.01 <- aov(ranges~1+regions, data=rangedata)
anova(lm.01)
```

That looks good! But are the data suitable for an anova in the first place? Let's test some assumptions.
```{r assumptions-test}
#Test some basic properties of the data
skewness(rangedata$ranges)
sd(rangedata$ranges)
mean(rangedata$ranges)

#Test the data to see if they fit the assumptions of an anova
shapiro.test(residuals(lm.01))
#leveneTest(rangedata$ranges~rangedata$regions,kruskal.test=TRUE)
fligner.test(rangedata$ranges~rangedata$regions)
bartlett.test(rangedata$ranges~rangedata$regions)
oneway.test(rangedata$ranges~1+rangedata$regions)
kruskal.test(rangedata$ranges~1+rangedata$regions)
```

These tests indicate that there is a slight right skew to the data, that it is non-normally distributed (Shapiro-Wilk's test) and non-homogeneous (Levene's Test). So, we need a multiple-comparisons method that is more robust to non-heteroscedasticity (while controlling the family-wise error rate). Luckily, there's a package for that:
```{r glht, warning=FALSE}
multicomp <- glht(lm.01, linfct = mcp(regions = "Tukey"), vcov = vcovHC)
summary(multicomp)
```

What do we find? Pitch-class imprecision was (a) lower in the Middle East-North Africa region than all other regions (p < 0.001), except Central Asia; (b) lower in Central Asia than the Circumpolar and Insular Pacific regions (p < 0.001), as well as East Asia and South America (p < 0.01); (c) higher in the Insular Pacific region than Europe (p < 0.01) and North America (p < 0.05); and (d) higher in the Circumpolar region than Europe (p < 0.05).  

## Plotting the data
To make a nice colourful figure, I want to plot the mean imprecision of each region with standard deviation error bars.

```{r plotting, warning=FALSE, fig.height=5, fig.width=10, fig.align='center'}
#Get the means 
rangemeans <- aggregate(rangedata$ranges, list(rangedata$regions), FUN=mean)
rangemeans$Group.1 <- factor(rangemeans$Group.1, levels = regionlevels)
rangemeans <- rangemeans[order(rangemeans$Group.1), ]
#Get the standard deviations
rangesds <- aggregate(rangedata$ranges, list(rangedata$regions), FUN=sd)
rangesds$Group.1 <- factor(rangesds$Group.1, levels = regionlevels)
rangesds <- rangesds[order(rangesds$Group.1), ]
#Make one dataframe
rdf <- data.frame(rangemeans$Group.1, rangemeans$x, rangesds$x)
names(rdf) <- c('regions','means','sds')
#Assign some plot parameters
cols <- c("#0C5BB0FF","#EE0011FF","#15983DFF","#149BEDFF","#FA6B09FF","#EC579AFF","#A1C720FF","#FEC10BFF","#9A703EFF","#16A08CFF")
regionlevelsshort <- c('Africa','Australia','Central Asia','Circumpolar','East Asia','Europe','Insular Pacific','Middle East','North America','South America')

#Make the plot
plot1 <- ggplot(rdf, aes(x=regions, y=means, fill=regions)) + 
  geom_bar(position=position_dodge(), stat="identity",
           colour = "black", size=0.3)+
  geom_errorbar(aes(ymin=means-sds, ymax=means+sds),
                size=0.3,
                width=0.2,
                position=position_dodge(0.9))+
  scale_fill_manual(values=cols)+
  scale_y_continuous(breaks=seq(0,350,50))+
  ylab("Pitch class imprecision (cents)")+
  theme_bw()+
  theme(text = element_text(size = 16), axis.text.y=element_text(size=14), axis.title.x = element_blank(),axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.line = element_blank(), axis.ticks.length.x = unit(0, "pt"), legend.position="top") 
show(plot1)

```

## What next
Now, can you try doing the same analysis with the other dataset in this package, pitchclass_inacc?
